---
title: "1D Inverse problem class activity"
author: "Drew Remmenga"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Setup

```{r}
library(viridis)
source("naturalSplineBasis.R")
source("naturalCubicSplineR.R")
source("InverseExampleFunctions.R")


suppressMessages(library( fields))
suppressMessages(library( fda))
```

# Create true function

True function is evaluated on a grid 1 to 200

```{r}
set.seed(123)

M<- 200
sGrid<- 1:M


uGrid<- sGrid/ M
truef<- 9*uGrid*(1-uGrid)^3
```

Now generate the observations via the W matrix 80 observations 

```{r}
N<- 80
W<- matrix(0,N,M)
I<- matrix( NA, N,2)
center<- rep(NA, N)
set.seed(222)
for(  k in 1:N){
c0<- sample( 1:M, 1, replace=FALSE)
w0<- sample( 5:(M/2), 1, replace=FALSE)
i<-  c( max( c(c0-w0,1) ) , min(c(c0+w0,M) ))
I[k,]<- i
W[k, i[1]:i[2] ] <- 1/(i[2]- i[1])
center[k]<- (i[2] + i[1])/2
}

```

Stats on W matrix just to check distribution of widths. 

```{r}

dim( W)

WInd<- W>0
stats(  rowSums( WInd))

```



Find expected value of the observations and create data.
The plot below emphasizes the problem of trying to look at these data without using the W matrix correctly. 
Here the expected observation without the error added is plotted against the center of the interval. Note
that without error we should see a smooth curve. The variation here is due the different sizes
of the weight function applied at different locations.  

```{r}
EY<- W%*%truef
fields.style()
plot( center, EY, pch=16)
set.seed( 222)
error<- .05*rnorm( N)
z<- EY +error
lines( sGrid, truef, col="blue")
title("expected value of observations and interval centers
      true curve (green)")

```

# Functions for fitting

The function **fitWBS**  fits a bspline to the the data correctly using the W weight matrix.  

Example of fitting with 6 knots

```{r}
fHat<- fitWBS(z,W,6,sGrid)
fields.style()
plot( sGrid, fHat, type="l")
lines( sGrid, truef, col=2)
```


The function **fitWBS**  fits a smoothing spline to the the data correctly using the W weight matrix. Here
the smoothing parameter lambda takes over the role of choosing the number of knots. In fact this function
simply puts a knot at every point of **sGrid** and relies on the roughness matrix/penalty and lambda to give
a good fit. 


Example fitting with lambda value  equal to 50. (This is too small!)

```{r}
fHat<- fitWSS(z,W, 50 ,sGrid)
fields.style()
plot( sGrid, fHat, type="l")
lines( sGrid, truef, col=2)
```


# Task 1

- How do you find the predicted values for the observations based on these estimates?
Use sgrid and fitwss.
- How would you find residuals from the fit? 
fhat - fitwbs
# Task 2
For the B spline model vary the knots from 4 to 15. 
and use 10/90 cross validation to estimate the best number of knots. 
```{r}
smp_size <- floor(0.9 * 200)
train_ind <- sample(200, size = smp_size)

train <- z[train_ind]
test <- z[-train_ind]
set.seed(123)

M<- 180
sGrid<- 1:M


uGrid<- sGrid/ M
truef<- 9*uGrid*(1-uGrid)^3
N<- 80
W<- matrix(0,N,M)
I<- matrix( NA, N,2)
center<- rep(NA, N)
set.seed(222)
for(  k in 1:N){
c0<- sample( 1:M, 1, replace=FALSE)
w0<- sample( 5:(M/2), 1, replace=FALSE)
i<-  c( max( c(c0-w0,1) ) , min(c(c0+w0,M) ))
I[k,]<- i
W[k, i[1]:i[2] ] <- 1/(i[2]- i[1])
center[k]<- (i[2] + i[1])/2
}
EY<- W%*%truef
set.seed( 222)
error<- .05*rnorm( N)
z<- EY +error
val = 4:15
crossvalidation = c()
for (x in val){
fHat<- fitWBS(z,W,x,sGrid)
fields.style()
plot( sGrid, fHat, type="l")
lines( sGrid, truef, col=2)
append(crossvalidation,val=mean(fHat-truef)/x)
}
```
# Task 3 
For the smoothing spline model use 10/90 cross validation to estimate lambda. 


```{r}
val= 50:150
for (x in val){
fHat<- fitWSS(z,W, x ,sGrid)
fields.style()
plot( sGrid, fHat, type="l")
}
lines( sGrid, truef, col=2)
```






